---
title: "Course project - Practical Machine Learning"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Introduction

The goal of this project is to predict the manner in which the 6 participants did their exercices. There are 5 different classes. We will divide the pml-training.csv in a training set and a test set to validate the model. With the pml-test.csv, we will predict the classe for each action.

```{r}
library(caret)
library(randomForest)
```



## Load the data


```{r}
training <- read.csv("pml-training.csv")
testing_class <- read.csv("pml-testing.csv")
```


```{r}
str(training)
```


# Clean data

Get rid of the variables where there is a lot of NAs.

```{r}
training2 <- training[,-c(grep("^kurtosis_",names(training)),
                          grep("^skewness_",names(training)),
                          grep("^var_",names(training)),
                          grep("^max_",names(training)),
                          grep("^min_",names(training)),
                          grep("^amplitude_",names(training)),
                          grep("^avg_",names(training)),
                          grep("^stddev_",names(training)), 
                          grep("^raw_timestamp_part_",names(training)),
                          grep("^cvtd_timestamp",names(training)),
                          grep("^new_window",names(training)) )]


dim(training2)
```





# Separate the training set in two groups

We want a training set (60% of the observations in training2) and a testing set (40% of the observations in training2).

```{r}
Intrainset <- createDataPartition(training$X,p=0.6,list=FALSE)

trainingset <- training2[Intrainset,]

testset <- training2[-Intrainset,]

dim(trainingset)
dim(testset)
```


We have 11774 observations in the training set. This is the set we will use to fit the model. We have 7848 observations in the test set.



# Fit a model

We will use the Random Forest to create classes. We will perform a PCA to reduce the number of predictors.

```{r, warning=FALSE}
trainrf <- train(classe~.-X,data=trainingset, preProcess="pca",method="rf")
predrf <- predict(trainrf, testset)

confusionMatrix(predrf, testset$classe)
```




We have an Accuracy of 0.976. 



# Make prediction with the validation data

We will run the model on the validation data to predict the classes.


```{r}

validate <- predict(trainrf, testing_class)
validate
```


We obtain the class for each of the observations in the validation set.


